On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-08-13 paper.tex
Writing introduction

\addbibresource{paper.bib}
\renewbibmacro*{volume+number+eid}{\ifentrytype{article}{\- \iffieldundef{volume}{}{Vol.~\printfield{volume},}\iffieldundef{number}{}{ No.~\printfield{number},}}}
\renewbibmacro{in:}{\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}
\newbibmacro{string+doi}[1]{\iffieldundef{doi}{\iffieldundef{url}{#1}{\href{\thefield{url}}{#1}}}{\href{http://dx.doi.org/\thefield{doi}}{#1}}}
\DeclareFieldFormat[article, inproceedings, inbook, book, thesis]{title}{\usebibmacro{string+doi}{\mkbibquote{#1}}}
\renewcommand*{\bibfont}{\footnotesize}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\T}{T}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\rk}{rk}

@@

The aim of this note is to show that $\K_2$-analogue of Horrocks theorem holds for the unstable group $K_{2, 4}(R)$.
From this we deduce an improved version of Tulenbayev's theorem on early stability for linear $K_2$-groups.

Let $R$ be a commutative ring and $\St(n, R)$ be the Steinberg group of rank $n$.
Denote by $\K_2(n, R)$ (resp. $\K_1(n, R)$) the kernel (resp. cokernel) of the canonical map $\phi\colon \St(n, R) \to \GL(n, R)$.
Recall that in~\cite[Theorem~5.1]{Tu83} M.~Tulenbayev proves the following result.
\begin{thm*} For $n \geq 5$ the Steinberg group functor $\St(n, -)$ satisfies $\mathbb{P}^1$-glueing, i.\,e. for every
 local ring $R$ the following square is pullback:
 \[\xymatrix{ \St(n, R)    \ar[r]\ar[d] & \ar[d] \St(n, R[t^{-1}] \\ 
              \St(\Phi, R[t]) \ar[r]       &        \St(n, R[t, t^{-1}]).}\]
\end{thm*}
The above theorem should be considered as a ``$\K_2$-analogue'' of Horrocks theorem on projective modules (see e.\,g.~\cite[Ch.~IV]{Lam10}).
Recall that this theorem asserts that $\mathbb{P}^1$-glueing holds for the functor $\mathrm{VB}_n$ 
 associating to each commutative ring $R$ the set of rank $n$ vector bundles over $R$.
On the level of $\K_1$ the analogous statement was first obtained by A.~Suslin for the functor $\K_1(n, R)$ under the assumption $n\geq 3$, see~\cite{Su77}.

A lot of effort has been put afterwards into generalizing Suslin's results to other groups (see e.\,g.~\cite{Abe83}). 
One of the more recent results in this direction is~\cite[Theorem~1.1]{St-poly} due to A.~Stavrova, in which
 $\mathbb{P}^1$-glueing for $K_1$ is proved for a large class of isotropic reductive groups.

>>>>

\addbibresource{paper.bib}
\renewbibmacro*{volume+number+eid}{\ifentrytype{article}{\- \iffieldundef{volume}{}{Vol.~\printfield{volume},}\iffieldundef{number}{}{ No.~\printfield{number},}}}
\renewbibmacro{in:}{\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}
\newbibmacro{string+doi}[1]{\iffieldundef{doi}{\iffieldundef{url}{#1}{\href{\thefield{url}}{#1}}}{\href{http://dx.doi.org/\thefield{doi}}{#1}}}
\DeclareFieldFormat[article, inproceedings, inbook, book, thesis]{title}{\usebibmacro{string+doi}{\mkbibquote{#1}}}
\renewcommand*{\bibfont}{\footnotesize}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\T}{T}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\rk}{rk}
\newcommand{\catname}[1]{{\normalfont\textbf{#1}}}

@@

The aim of this note is to show that $\K_2$-analogue of Horrocks theorem holds for the unstable group $K_{2}(4, R)$.
From this theorem deduce an improved version of Tulenbayev's theorem on early stability for linear $K_2$-groups.

Let $R$ be a commutative ring and $\St(n, R)$ be the Steinberg group of rank $n$.
Denote by $\K_2(n, R)$ (resp. $\K_1(n, R)$) the kernel (resp. cokernel) of the canonical map $\phi\colon \St(n, R) \to \GL(n, R)$.
Recall that in~\cite[Theorem~5.1]{Tu83} M.~Tulenbayev proves the following result.
%TODO: Maybe formulate the final result, i. e. early stability theorem
\begin{thm*} For $n \geq 5$ the Steinberg group functor $\St(n, -)$ satisfies $\mathbb{P}^1$-glueing, i.\,e. for every
 local ring $R$ the following square is pullback:
 \[\xymatrix{ \St(n, R)    \ar[r]\ar[d] & \ar[d] \St(n, R[t^{-1}]) \\ 
              \St(\Phi, R[t]) \ar[r]       &        \St(n, R[t, t^{-1}]).}\]
\end{thm*}
It is not hard to see that $\mathbb{P}^1$-glueing property holds simulteneously for Steinberg group functor $\St(n, -)$ and the functor $\K_2(n, -)$. Therefore the above theorem should be considered as a ``$\K_2$-analogue'' of the classical Horrocks theorem on projective modules (see e.\,g.~\cite[Ch.~IV]{Lam10}).
Recall that the latter result asserts that $\mathbb{P}^1$-glueing holds for the functor $\mathrm{VB}_n$ associating to each commutative ring $R$ the set of rank $n$ vector bundles over $R$ (the set $\mathrm{VB}_n$ should be thus considered as the unstable analogue of $K_0$-functor).

The corresponding result on $\mathbb{P}^1$-glueing for for the functor $\K_1(n, R)$,  $n\geq 3$ was first obtained by A.~Suslin as one of the two ingredients in the proof of the $\K_1$-analogue of Serre problem, see~\cite{Su77}. A lot of effort has been put afterwards into generalizing Suslin's results to other groups (see e.\,g.~\cite{Abe83}).  One of the more recent results in this direction is~\cite[Theorem~1.1]{St-poly} due to A.~Stavrova, in which $\mathbb{P}^1$-glueing property is proved for functors $\K_1^G$ modeled on a large class of isotropic reductive groups (consisting, roughly speaking, of groups having local isotropic rank at least $2$). %TODO: Better phrasing

Let $\mathcal{F}\colon \catname{CRings} \to \catname{Sets}_*$ be a functor from commutative rings to pointed sets. We say that an analogue of Serre problem holds for $\mathcal{F}$ if for any field $k$ there is an isomorphism $\mathcal{F}(k[t_1, \ldots, t_n]) = \mathcal{F}(k)$. The well-known counterexample of P.~Cohn shows that the $\K_1$-analogue of Serre problem fails for $\K_1(2, -)$, see~\cite[\S~I.8]{Lam10}. Similarly, it is known that an analogue of Serre problem fails for $\K_2(3, -)$, see~\cite{We14}. Thus one might ask whether an analogue of Serre problem holds for $\K_2(4, -)$, the only case not covered by Tulenbayev's theorem.

Typically, the proof of an analogue of Serre problem for a functor $\mathcal{F}$ (e.\,g. $\mathrm{VB}_n$, $K_1^G$, $K_2(n, -)$) reduces the following two steps: 
\begin{enumerate}
 \item proving that an analogue of local-global principle holds for $\mathcal{F}$ (see ??? for the definition below);
 \item proving that $\mathcal{F}$ satisfies $\mathbb{P}^1$-glueing.
\end{enumerate}

In~\cite{LS17} A.~Lavrenov and the author showed that the local-global principle holds for $\St(4, -)$.

Our first main result shows that it is indeed so.
\begin{thm}
 The functor $\St(4, -)$ satisfies $\mathbb{P}^1$-glueing.
\end{thm}
Consequently, there is an isomorphism $\K_2(4, k[t_1, \ldots, t_n]) \cong \K_2(4, k)$
To deduce the local-global principle for $K_2(4, -)$ which has already been obtained in~\cite{LS17} this
